Smart Drill-Down

### 2014-11-30 TableSummarization/Paper/algorithms.tex
small fixes

As a first step towards solving Problem~\ref{prob:optimal-rule-list}, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight. Note that the weight of a rule is independent of its $MCount$. The $MCount$ of a rule is the number of tuples that have been `assigned' to it, and each tuple assigned to rule $r$ contributes $W(r)$ to the total score. Thus if the rules are not in decreasing order by weight in a rule list $R$, then switching the order of rules in $R$ transfers some tuples from a lower weight rule to a higher weight rule, which can increase total score.

@@

\noindent Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a \emph{set} of rules as follows:

\begin{definition}\label{def:set-score}
Let $R$ be a set of rules. Then the Score of $R$ is
$$\text{Score}(R) = \text{Score}(R^{\prime})$$
where $R^{\prime}$ is the list of rules obtained by ordering the rules in $R$ in decreasing order by weight.
\end{definition}

@@

\subsection{Algorithm Overview}\label{sec:alg-overview}
Given that the problem is {\sc NP-Hard}, we now present our algorithms 
for approximating the solution to Problem~\ref{prob:optimal-rule-set}. 
The problem consists of finding a set of given size $k$, that maximizes 
Score.

@@

\subsection{Greedy Approximation Algorithm}\label{sec:greedy-approx}
\stitle{Submodularity:} We will now show that the Score function over sets of rules has a property called {\em submodularity}, giving us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy fashion.

@@

In order to avoid counting too many rules, we use an idea from the {\em a-priori} algorithm for frequent itemset mining~\cite{apriori}. Recall that the a-priori algorithm is used to find all frequent itemsets that have a support greater than a threshold. Unlike the frequent itemset mining algorithm, our goal is to find the single best marginal rule. Since we only aim to find one rule at a time, our pruning power is significantly higher than a vanilla a-priori algorithm, and we terminate in much fewer passes over the dataset. 
We compute the best marginal rule over several passes, with the maximum number of passes equal to the maximum size of a rule. In the $j^{th}$ pass, we compute counts and marginal values for rules of size $j$. To give an example, suppose we had three columns $c_1$, $c_2$, and $c_3$. In the first pass, we would compute the counts and marginal values of all rules of size $1$. In the second pass, instead of finding marginal values for all size $2$ rules, we can use our knowledge of counts from the first pass to upper bound the potential counts and marginal values of size $2$ rules, and be more selective about which rules to count in the second pass. For instance, suppose we know that the rule $(a, \star, \star)$ has a count of $1000$, while $(\star, b, \star)$ has a count of $100$. Then for any value $c$ in column $c_3$ we would know that the count of $(\star, b, c)$ is at most $100$ because it cannot exceed that of $(\star, b, \star)$. This implies that the maximum marginal value of any super-rule of $(\star, b, c)$ having weight $\leq m_w$ is at most $100m_w$. If the rule $(a, \star, \star)$ has a marginal value of $800$, then the marginal value of any super-rule of $(\star, b, \star)$ cannot possibly exceed that of $(a, \star, \star)$. Since our aim is to only find the highest marginal value rule, we can skip counting for all super-rules of $(\star, b, \star)$ for future passes.

>>>>

As a first step towards solving Problem~\ref{prob:optimal-rule-list}, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight. Note that the weight of a rule is independent of its $MCount$. The $MCount$ of a rule is the number of tuples that have been `assigned' to it, and each tuple assigned to rule $r$ contributes $W(r)$ to the total score. Thus, if the rules are not in decreasing order by weight in a rule list $R$, then switching the order of rules in $R$ transfers some tuples from a lower weight rule to a higher weight rule, which can increase total score.

@@

\noindent Thus, it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a \emph{set} of rules as follows:

\begin{definition}\label{def:set-score}
Let $R$ be a set of rules. Then the Score of $R$ is
$$\text{Score}(R) = \text{Score}(R^{\prime})$$
where $R^{\prime}$ is the list of rules obtained by ordering the rules in the set $R$ in decreasing order by weight.
\end{definition}

@@

\subsection{Algorithm Overview}\label{sec:alg-overview}
Given that the problem is {\sc NP-Hard}, we now present our algorithms 
for approximating the solution to Problem~\ref{prob:optimal-rule-set}. 
The problem consists of finding a set of rules, given size $k$, that maximizes 
Score.

@@

\subsection{Greedy Approximation Algorithm}\label{sec:greedy-approx}
\stitle{Submodularity:} We will now show that the Score function over sets of rules has a property called {\em submodularity}, giving us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$: $f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, it is well known that the solution to the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy fashion.

@@

In order to avoid counting too many rules, we leverage a technique inspired by the {\em a-priori} algorithm for frequent itemset mining~\cite{apriori}. Recall that the a-priori algorithm is used to find all frequent itemsets that have a support greater than a threshold. Unlike the a-priori algorithm, our goal is to find the single best marginal rule. Since we only aim to find one rule at a time, our pruning power is significantly higher than a vanilla a-priori algorithm, and we terminate in much fewer passes over the dataset.

We compute the best marginal rule over multiple passes on the dataset, with the maximum number of passes equal to the maximum size of a rule. In the $j^{th}$ pass, we compute counts and marginal values for rules of size $j$. To give an example, suppose we had three columns $c_1$, $c_2$, and $c_3$. In the first pass, we would compute the counts and marginal values of all rules of size $1$. In the second pass, instead of finding marginal values for all size $2$ rules, we can use our knowledge of counts from the first pass to upper bound the potential counts and marginal values of size $2$ rules, and be more selective about which rules to count in the second pass. For instance, suppose we know that the rule $(a, \star, \star)$ has a count of $1000$, while $(\star, b, \star)$ has a count of $100$. Then for any value $c$ in column $c_3$ we would know that the count of $(\star, b, c)$ is at most $100$ because it cannot exceed that of $(\star, b, \star)$. This implies that the maximum marginal value of any super-rule of $(\star, b, c)$ having weight $\leq m_w$ is at most $100m_w$. If the rule $(a, \star, \star)$ has a marginal value of $800$, then the marginal value of any super-rule of $(\star, b, \star)$ cannot possibly exceed that of $(a, \star, \star)$. Since our aim is to only find the highest marginal value rule, we can skip counting for all super-rules of $(\star, b, \star)$ for future passes.

### 2014-11-30 TableSummarization/Paper/sampling.tex
small fixes

\subsection{Dynamic Sampling for Large Tables}\label{sec:sampling}
We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory. First in Section~\ref{sec:sample-using}, we motivate the need for sampling. Then we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints. The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next. Then in Section~\ref{sec:sampling_algorithms}, we describe ways to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk. 
\papertext{In the technical report,}
\techreporttext{Finally,}
we mention some additional optimizations we can make, and
describe how we can set the minimum sample size required from the SampleHandler.

@@

Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$. Thus the $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$. The SampleHandler also computes $N_s$ when a sample is created. Then we run Algorithm~\ref{algo:best-rule-set} on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display. The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table. In addition since the sample is uniformly random, we can also compute confidence intervals on the estimated count of each displayed rule, although we do not currently display the confidence intervals.

@@

At any stage, we have a tree $U$ displayed to the user, with each node of the tree corresponding to a displayed rule. The tree is formed as follows: The root of the tree corresponds to the trivial rule. And suppose the user expands a node with rule $r$, resulting in rules $r_1, r_2, .. r_k$ being displayed. Then we add children nodes to the expanded node, corresponding to rules $r_1, r_2, .. r_k$. Note that multiple nodes of the tree may correspond to the same rule. Thus even though the structure displayed to the user is a tree, the set of rules displayed forms a partially ordered set (poset) using the sub-rule ordering. Internal nodes of $U$ are ones that have been expanded (drilled down on), while leaves are nodes that have not been expanded. Let $L$ be the set of leaves. Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for rules corresponding to leaf nodes.
In the rest of this section, we abuse notation to use the term rule to refer to any node that corresponds to that rule.

@@

We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$, without effectively changing the problem. For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$. Tree $U$ has $m$ special nodes $r_1, r_2, ... r_m$, and each $r_i$ has two children $r_{i,1}, r_{i,2}$. These $2m$ children are all leaves, and all leaves other than these have expansion probability $0$. 
The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$. In reality, the $S$ values cannot be exactly zero when the first argument is the trivial rule, but we can make it small enough such that any optimal solution will set $n_r = 0$ when $r$ is not a special node or its child. Thus each each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else. Thus $ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j}) \forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace$. In addition, $S(r_i, r_{i,1}) = 1$ (again, it cannot be exactly $1$, but can be brought arbitrarily close to $1$), and $S(r_i, r_{i,2}) = 1 - w_i$. Finally, for each i, $p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$. Thus each individual $p_{r_{i,1}}$ value is higher than all $p_{r_{i,2}}$ values combined, and $M$ is high enough to cover all $r_{i,1}$. Thus in any optimal solution, $ess(r_{i,1})$ must be $= minSS$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$. For all $i$, we must have either $n_{r_i} = n_{r_{i,2}} = 0 \land n_{r_{i,1}} = minSS$ (iff $ess(r_{i,2}) < minSS$) or $n_{r_i} = minSS \land n_{r_{i,1}} = 0 \land n_{r_{i,2}} = minSS (1 - S(r_i, r_{i,2})) = minSS \times w_i$ (iff $ess(r_{i,2}) = minSS$). The latter option consumes $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value. Thus having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (as it consumes additional memory $\propto w_i$ and gives additional probability value $\propto v_i$). Moreover the additional memory available (on top of the $m \times minSS$ required to cover all $r_{i,1}$s) is $W\times minSS$. Hence solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem. 
\end{proof}

@@

Now consider a rule in $r_0 \in U \setminus L$ along with all its children. Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children. Then by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$. This allows us to effectively split the problem into multiple subproblems,  with one subproblem per $M_{r_0}$. Thus for each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$. Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{ess(r) \geq minSS}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$. Then we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s. We describe both these steps in detail below:

Let $r_0 \in U \setminus L$. Let $d$ be the number of leaf children of $r_0$. Let the children be $r_1, r_2, ... r_d$. For any child $r_i$, $n_{r_i}$ only contributes to it's own $ess$, whereas $n_{r_0}$ contributes to the $ess$ of all children $r_1, ... r_d$. Given a value of $n_{r_0}$, in a locally optimal solution, each child $r_i$ must satisfy:
\squishlist
\item If $n_{r_0}S(r_0, r_i) \geq minSS$, then $n_{r_i} = 0$ because otherwise, decreasing $n_{r_i}$ to $0$ would lower its sampling cost without improving its probability score. 
\item If $n_{r_0}S(r_0, r_i) < minSS$, then either $n_{r_i} = 0$ or $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$. This is because if $n_{r_i}$ is between $0$ and $minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $0$, and if it is $> minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $minSS - n_{r_0}S(r_0, r_i)$. Both these decreases would decrease sampling cost without affecting probability score. 
\squishend
Thus there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess \geq minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$. There are $3^d$ ways to assign each child to one of these categories, and each of those potentially gives us one locally optimal solution. Consider any such locally optimal solution $e$. For $e$ let children $r_{i_1}, r_{i_2}, ... r_{i_m}$ be in the first category, $r_{i_{m+1}}, .. r_{i_M}$ be in the second category, and $r_{i_{M+1}}, .. r_{i_d}$ in the third.  Then the `probability value' of solution $e$ is given by : 
$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
and its `Sampling Cost' is
$$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
Thus there are at most $3^d$ locally optimal solutions. $d$ is usually small (it is at most $k$, the size of the rule-list we create on every user click), even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.

\begin{comment}
Let $r_0 \in U \setminus L$. Let $d$ be the number of leaf children of $r_0$. Let the children be $r_1, r_2, ... r_d$, ordered such that $S(r_0, r_i) > S(r_0, r_j) \forall i > j$. Then we can show that for any locally optimal solution, there exist integers $i_m, i_M$ such that the following conditions hold:
$$1 \leq i_m \leq i_M \leq d$$
$$n_{r_0} = \frac{minSS}{S(r_0, r_{i_m})}$$
$$1 \leq j \leq i_m \Rightarrow n_{r_j} = 0$$
$$i_m < j \leq i_M \Leftrightarrow n_{r_j} = minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
$$i_M < j \Rightarrow n_{r_j} = 0$$
And thus
$$ess(r_j) \geq minSS \Leftrightarrow j \leq i_M$$

@@

In addition, in order to express our problem as a convex minimization problem, we negate the objective function and aim to minimize it (which is equivalent to maximizing the original objective function). Thus our new optimization problem becomes
\begin{problem}\label{prob:sample-sizes-hinge-loss}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose a real number $n_r \geq 0$ for each $r \in U$ so as to $\textrm{minimize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{max}(-1, -\frac{ess(r^{\prime})}{minSS} $$
subject to :
$$\sum_{r \in U} n_r \leq M$$
\end{problem}
The constraint is linear in the $n_r$ variables, and hence convex. Each $ess$ value is a linear function of the $n_r$s, which makes $-\frac{ess(r^{\prime})}{minSS}$ convex. The constant function $-1$ is convex as well. Since the maximum of two convex functions is convex, Problem~\ref{prob:sample-sizes-hinge-loss} is a convex minimization problem, which means that its local optimum is also its global optimum. Thus we can initialize all $n_r$s to $0$ and then use stochastic gradient descent (or any other local optimization technique) to find their optimum values.

@@

Thus a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$. This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function (where weight of a rule equals number of non-$\star$ values of the rule). Let $c$ be the column with the fewest distinct values. Say it has $|c|$ values. Then the rule that has the most frequent value of $c$, and $\star$ everywhere else, must have a score of at least $\frac{|T|}{|c|}$. For example, if the table has $10000$ tuples in all, and there is a `Education' column that has 5 possible values, then the most frequent value of Education must occur at least $2000$ times. So the rule with the most frequent value for Education, and $\star$s elsewhere, must have a score of at least $2000$.

>>>>

\subsection{Dynamic Sampling for Large Tables}\label{sec:sampling}
We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory. First in Section~\ref{sec:sample-using}, we motivate the need for sampling. Then we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints. The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next. Then in Section~\ref{sec:sampling_algorithms}, we describe ways to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk. 
\papertext{In the technical report,}
\techreporttext{Finally,}
we mention some additional optimizations we can make, and
describe how we can set the minimum sample size required from the SampleHandler.
\agp{revisit}

@@

Whenever the user drills down on a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$. Thus, the $T_s$ of the returned sample consists of a uniformly random set of tuples covered by $r$. The SampleHandler also computes $N_s$ when a sample is created. Then we run Algorithm~\ref{algo:best-rule-set} on sample $s$ (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display. The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get estimated counts on the entire table. In addition since the sample is uniformly random, we can also compute confidence intervals on the estimated count of each displayed rule, although we do not currently display the confidence intervals.

@@

At any stage, we have a tree $U$ displayed to the user, with each node of the tree corresponding to a displayed rule. The tree is formed as follows: The root of the tree corresponds to the trivial rule. And suppose the user expands a node with rule $r$, resulting in rules $r_1, r_2, .. r_k$ being displayed. Then we add children nodes to the expanded node, corresponding to rules $r_1, r_2, .. r_k$. Note that multiple nodes of the tree may correspond to the same rule. Thus, even though the structure displayed to the user is a tree, the set of rules displayed forms a partially ordered set (poset) using the sub-rule ordering. Internal nodes of $U$ are ones that have been expanded (drilled down on), while leaves are nodes that have not been expanded. Let $L$ be the set of leaves. Each leaf is something that the user can potentially expand in the next step, and thus we would like to have pre-fetched samples for rules corresponding to leaf nodes.
In the rest of this section, we abuse notation to use the term rule to refer to any node that corresponds to that rule.

@@

We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$, without effectively changing the problem. For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$. Tree $U$ has $m$ special nodes $r_1, r_2, ... r_m$, and each $r_i$ has two children $r_{i,1}, r_{i,2}$. These $2m$ children are all leaves, and all leaves other than these have expansion probability $0$. 
The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$. In reality, the $S$ values cannot be exactly zero when the first argument is the trivial rule, but we can make it small enough such that any optimal solution will set $n_r = 0$ when $r$ is not a special node or its child. Thus, each each leaf $r_{i,j}$ gets tuples either from its own sample, or the sample of its parent, and from nowhere else. Thus, $ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j}) \forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace$. In addition, $S(r_i, r_{i,1}) = 1$ (again, it cannot be exactly $1$, but can be brought arbitrarily close to $1$), and $S(r_i, r_{i,2}) = 1 - w_i$. Finally, for each i, $p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$. Thus, each individual $p_{r_{i,1}}$ value is higher than all $p_{r_{i,2}}$ values combined, and $M$ is high enough to cover all $r_{i,1}$. Thus, in any optimal solution, $ess(r_{i,1})$ must be $= minSS$, and we're left to decide which $i$'s should also have $ess(r_{i,2}) = minSS$. For all $i$, we must have either $n_{r_i} = n_{r_{i,2}} = 0 \land n_{r_{i,1}} = minSS$ (iff $ess(r_{i,2}) < minSS$) or $n_{r_i} = minSS \land n_{r_{i,1}} = 0 \land n_{r_{i,2}} = minSS (1 - S(r_i, r_{i,2})) = minSS \times w_i$ (iff $ess(r_{i,2}) = minSS$). The latter option consumes $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value. Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem (as it consumes additional memory $\propto w_i$ and gives additional probability value $\propto v_i$). Moreover the additional memory available (on top of the $m \times minSS$ required to cover all $r_{i,1}$s) is $W\times minSS$. Hence solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem. 
\end{proof}

@@

Now consider a rule in $r_0 \in U \setminus L$ along with all its children. Let $M_{r_0}$ denote the set containing $r_0$ and all its leaf children. Then by our simplification, the number of tuples $n_{r^{\prime}}$ for any rule in $r^{\prime} \in M_{r_0}$ only affects the ess value of rules in $M_{r_0}$. This allows us to effectively split the problem into multiple subproblems,  with one subproblem per $M_{r_0}$. Thus, for each non-leaf rule $r_0$ and all its children, we compute all `locally optimal' assignments of $n_r \mid r \in M_{r_0}$. Locally optimal means that we cannot get a higher value of `probability value' $\sum_{r \in M_{r_0}} p_rI_{ess(r) \geq minSS}$ for the same `sampling cost' $\sum_{r\in M_{r_0}} n_r$. Then we can use dynamic programming to combine the locally optimal solutions of different $M_{r_0}$s. We describe both these steps in detail below:

Let $r_0 \in U \setminus L$. Let $d$ be the number of leaf children of $r_0$. Let the children be $r_1, r_2, ... r_d$. For any child $r_i$, $n_{r_i}$ only contributes to it's own $ess$, whereas $n_{r_0}$ contributes to the $ess$ of all children $r_1, ... r_d$. Given a value of $n_{r_0}$, in a locally optimal solution, each child $r_i$ must satisfy:
\squishlist
\item If $n_{r_0}S(r_0, r_i) \geq minSS$, then $n_{r_i} = 0$ because otherwise, decreasing $n_{r_i}$ to $0$ would lower its sampling cost without improving its probability score. 
\item If $n_{r_0}S(r_0, r_i) < minSS$, then either $n_{r_i} = 0$ or $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$. This is because if $n_{r_i}$ is between $0$ and $minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $0$, and if it is $> minSS - n_{r_0}S(r_0, r_i)$, then we can decrease it to $minSS - n_{r_0}S(r_0, r_i)$. Both these decreases would decrease sampling cost without affecting probability score. 
\squishend
Thus, there are three kinds of children $r_i$: Those with $ess \geq minSS$ but $n_{r_i} = 0$, those with $ess < minSS$ and $n_{r_i} = 0$, and those with $ess \geq minSS$ and $n_{r_i} = minSS - n_{r_0}S(r_0, r_i)$. There are $3^d$ ways to assign each child to one of these categories, and each of those potentially gives us one locally optimal solution. Consider any such locally optimal solution $e$. For $e$ let children $r_{i_1}, r_{i_2}, ... r_{i_m}$ be in the first category, $r_{i_{m+1}}, .. r_{i_M}$ be in the second category, and $r_{i_{M+1}}, .. r_{i_d}$ in the third.  Then the `probability value' of solution $e$ is given by : 
$$P(e) = \sum_{j = 1}^{i_M}p_{j}$$
and its `Sampling Cost' is
$$S(e) = \frac{minSS}{S(r_0, r_{i_m})} + \sum_{j=i_m+1}^{i_M} minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
Thus, there are at most $3^d$ locally optimal solutions. $d$ is usually small (it is at most $k$, the size of the rule-list we create on every user click), even when the rule tree $U$ itself is big, and so we can enumerate all $3^d$ locally optimal solutions and find their sampling cost and probability scores.

\begin{comment}
Let $r_0 \in U \setminus L$. Let $d$ be the number of leaf children of $r_0$. Let the children be $r_1, r_2, ... r_d$, ordered such that $S(r_0, r_i) > S(r_0, r_j) \forall i > j$. Then we can show that for any locally optimal solution, there exist integers $i_m, i_M$ such that the following conditions hold:
$$1 \leq i_m \leq i_M \leq d$$
$$n_{r_0} = \frac{minSS}{S(r_0, r_{i_m})}$$
$$1 \leq j \leq i_m \Rightarrow n_{r_j} = 0$$
$$i_m < j \leq i_M \Leftrightarrow n_{r_j} = minSS - \frac{minSS}{S(r_0, r_{i_j})}$$
$$i_M < j \Rightarrow n_{r_j} = 0$$
And Thus,
$$ess(r_j) \geq minSS \Leftrightarrow j \leq i_M$$

@@

In addition, in order to express our problem as a convex minimization problem, we negate the objective function and aim to minimize it (which is equivalent to maximizing the original objective function). Thus, our new optimization problem becomes
\begin{problem}\label{prob:sample-sizes-hinge-loss}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose a real number $n_r \geq 0$ for each $r \in U$ so as to $\textrm{minimize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}\textrm{max}(-1, -\frac{ess(r^{\prime})}{minSS} $$
subject to :
$$\sum_{r \in U} n_r \leq M$$
\end{problem}
The constraint is linear in the $n_r$ variables, and hence convex. Each $ess$ value is a linear function of the $n_r$s, which makes $-\frac{ess(r^{\prime})}{minSS}$ convex. The constant function $-1$ is convex as well. Since the maximum of two convex functions is convex, Problem~\ref{prob:sample-sizes-hinge-loss} is a convex minimization problem, which means that its local optimum is also its global optimum. Thus, we can initialize all $n_r$s to $0$ and then use stochastic gradient descent (or any other local optimization technique) to find their optimum values.

@@

Thus, a reasonable value of minSS can be found by obtaining a bound on $\frac{1-x}{x}$. This is hard to do for arbitrary weighting functions, but we can do it for the Size weighting function (where weight of a rule equals number of non-$\star$ values of the rule). Let $c$ be the column with the fewest distinct values. Say it has $|c|$ values. Then the rule that has the most frequent value of $c$, and $\star$ everywhere else, must have a score of at least $\frac{|T|}{|c|}$. For example, if the table has $10000$ tuples in all, and there is a `Education' column that has 5 possible values, then the most frequent value of Education must occur at least $2000$ times. So the rule with the most frequent value for Education, and $\star$s elsewhere, must have a score of at least $2000$.
